Comparison of Dimensionality Reduction Schemes for Parallel Global Optimization Algorithms

### 2018-09-05 2018/RSD/MultiEvolvents/RSD.tex
RSD: fix style issues claimed by Springer's editors

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%%%%%
\usepackage{eqnarray,amsmath}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\setlength{\textfloatsep}{4pt plus 1.0pt minus 1.5pt}
\setlength{\floatsep}{4pt plus 1.0pt minus 1.5pt}

\begin{document}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
%
\title{Comparison of dimensionality reduction schemes for parallel global optimization
algorithms
\thanks{The study was supported by the Russian Science Foundation, project No 16-11-
10150}}
%
\titlerunning{Dimension reduction schemes in parallel GO}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov \and
Vladislav Sovrasov \and
Ilya Lebedev}
%
\authorrunning{K. Barkalov et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}\\
\email{sovrasov.vlad@gmail.com}\\
\email{ilya.lebedev@itmm.unn.ru}\\
\url{http://hpc-education.unn.ru/}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This work considers a parallel algorithms for solving multi-extremal optimization problems.
Algorithms are developed within the framework of the information-statistical approach and
implemented in a parallel solver Globalizer. The optimization problem is solved by reducing
the multidimensional problem to a set of joint one-dimensional problems that are solved in
parallel. Five types of Peano-type space-filling curves are employed to reduce dimension. The
results of computational experiments carried out on several hundred test problems are discussed.

@@

%------------------------------------------------------------------------------
\section{Methods of Dimension Reduction}
\subsection{Single evolvent}

@@

%------------------------------------------------------------------------------
\subsection{Shifted evolvents}
\label{sec:shifted}

@@

%------------------------------------------------------------------------------
\subsection{Rotated evolvents}
The application of the scheme for building the multiple evolvents (hereinafter called the shifted
evolvents or $S$-evolvents) described in Subsection \ref{sec:shifted} allows to preserve the
information on the nearness of the points in the multidimensional space and, therefore, to
provide more precise (as compared to a single evolvent) estimate of Lipschitz constant in the
search process. However, this approach has serious restrictions, which narrow the applicability
of the parallel algorithms, designed on the base of the $S$-evolvents (see the end of the section
\ref{sec:seq_comp}).

@@

%------------------------------------------------------------------------------
\subsection{Non-Univalent evolvent}
%\begin{Russian}

@@

%------------------------------------------------------------------------------
\subsection{Smooth evolvent}

@@

%------------------------------------------------------------------------------
\subsection{Parallel algorithm exploiting a set of evolvents}
\label{sec:parallel_evolvents}
Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the
problems
\[
\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
particular processor are interpreted as the results of the trials in the rest problems (in the
corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
[0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
operations.
\par
1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
\par
2. Inform the rest of processors about the start of the trial execution at the point \( y^k\) (the
blocking of the point \(y^k\) ).
\par
3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
\(x{}^{k_1} ,\dots,x{}^{k_s} \)
\par
4. Inform the rest of processors about the trial results at the point \(y^k\).
\par

@@

%------------------------------------------------------------------------------
\subsection{Comparison of the sequential evolvents}
\label{sec:seq_comp}
%\begin{Russian}
In order to understand whether any type of evolvents listed above has an essential advantage as
compared to other ones, the operating characteristics of the index method with different types
of evolvents have been obtained for the classes GKLS 2d Simple and GKLS 3d Simple. The
global minimum was considered to be found if the algorithm generates a trial point $y^k$ in the
$\delta$-vicinity of the global minimizer, i.e. $\left\|y^k-y^\ast\right\|_\infty\leq\delta$. The size
of the vicinity was selected as $\delta = 0.01\left\|b-a\right\|_\infty$. In case of GKLS
$\delta=0.01$.

@@

%------------------------------------------------------------------------------
\subsection{Parallel rotated evolvents}
\label{sec:results_parallel}
In order to evaluate the efficiency of the parallel algorithm from
Sec.~\ref{sec:parallel_evolvents}, the numerical experiments on the GKLS 4d (Hard, Simple)
classes and on the GKLS 5d (Hard, Simple) ones were conducted. The value of \(r\) in all
experiments was equal to 5.0, the size of the \(\delta\)-vicinity of the known solution was
increased up to 0.3. When solving the series of problems, up to 8 cluster nodes and up to 32
computational threads on each node were employed.

@@

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%%%%%
\usepackage{eqnarray,amsmath}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[misc]{ifsym}
\usepackage[labelsep=period]{caption}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\setlength{\textfloatsep}{4pt plus 1.0pt minus 1.5pt}
\setlength{\floatsep}{4pt plus 1.0pt minus 1.5pt}

\begin{document}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
%
\title{Comparison of Dimensionality Reduction Schemes for Parallel Global Optimization
Algorithms}
%
\titlerunning{Dimension Reduction Schemes in Parallel GO}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov (\Letter),
Vladislav Sovrasov and
Ilya Lebedev}
%
\authorrunning{K. Barkalov et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}\\
\email{sovrasov.vlad@gmail.com}\\
\email{ilya.lebedev@itmm.unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This work considers a parallel algorithms for solving multi-extremal optimization problems.
Algorithms are developed within the framework of the information-statistical approach and
implemented in a parallel solver Globalizer. The optimization problem is solved by reducing
the multidimensional problem to a set of joint one-dimensional problems that are solved in
parallel. Five types of Peano-type space-filling curves are employed to reduce dimension. The
results of computational experiments carried out on several hundred test problems are discussed.

@@

%------------------------------------------------------------------------------
\section{Methods of Dimension Reduction}
\subsection{Single Evolvent}

@@

%------------------------------------------------------------------------------
\subsection{Shifted Evolvents}
\label{sec:shifted}

@@

%------------------------------------------------------------------------------
\subsection{Rotated Evolvents}
The application of the scheme for building the multiple evolvents (hereinafter called the shifted
evolvents or $S$-evolvents) described in Subsection \ref{sec:shifted} allows to preserve the
information on the nearness of the points in the multidimensional space and, therefore, to
provide more precise (as compared to a single evolvent) estimate of Lipschitz constant in the
search process. However, this approach has serious restrictions, which narrow the applicability
of the parallel algorithms, designed on the base of the $S$-evolvents (see the end of the section
\ref{sec:seq_comp}).

@@

%------------------------------------------------------------------------------
\subsection{Non-Univalent Evolvent}
%\begin{Russian}

@@

%------------------------------------------------------------------------------
\subsection{Smooth Evolvent}

@@

%------------------------------------------------------------------------------
\subsection{Parallel Algorithm Exploiting a Set of Evolvents}
\label{sec:parallel_evolvents}
Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the
problems
\[
\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
particular processor are interpreted as the results of the trials in the rest problems (in the
corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
[0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
operations.
\par
1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
\par
2. Inform the rest of processors about the start of the trial execution at the point \( y^k\) (the
blocking of the point \(y^k\) ).
\par
3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
\(x{}^{k_1} ,\dots,x{}^{k_s} \)
\par
4. Inform the rest of processors about the trial results at the point \(y^k\).
\par

@@

%------------------------------------------------------------------------------
\subsection{Comparison of the Sequential Evolvents}
\label{sec:seq_comp}
%\begin{Russian}
In order to understand whether any type of evolvents listed above has an essential advantage as
compared to other ones, the operating characteristics of the index method with different types
of evolvents have been obtained for the classes GKLS 2d Simple and GKLS 3d Simple. The
global minimum was considered to be found if the algorithm generates a trial point $y^k$ in the
$\delta$-vicinity of the global minimizer, i.e. $\left\|y^k-y^\ast\right\|_\infty\leq\delta$. The size
of the vicinity was selected as $\delta = 0.01\left\|b-a\right\|_\infty$. In case of GKLS
$\delta=0.01$.

@@

%------------------------------------------------------------------------------
\subsection{Parallel Rotated Evolvents}
\label{sec:results_parallel}
In order to evaluate the efficiency of the parallel algorithm from
Sec.~\ref{sec:parallel_evolvents}, the numerical experiments on the GKLS 4d (Hard, Simple)
classes and on the GKLS 5d (Hard, Simple) ones were conducted. The value of \(r\) in all
experiments was equal to 5.0, the size of the \(\delta\)-vicinity of the known solution was
increased up to 0.3. When solving the series of problems, up to 8 cluster nodes and up to 32
computational threads on each node were employed.

@@

\section*{Acknowledgements}
The study was supported by the Russian Science Foundation, project No 16-11-
10150
